---
title: "Airline IBM Regression Project"
output: html_notebook
---

# Airline IBM Regression Project

## Introduction

This project will include Exploratory data analysis and classification models to analyze the given data and attributes that can predict whether or not a plane will be delayed or on time.

To start this project, I will load in a couple of libraries as well via a Github URL and the "read.csv" function. Since the .csv file was too large to be placed directly into GitHub, I had to use Git LFS to push the file through git into my repository as a prerequisite.

```{r}
library(tidyverse)
library(ggplot2)
```

```{r}
url<- "https://media.githubusercontent.com/media/Wesnahika/Airline-IBM-Regression-Project/main/airline_2m.csv"
airline <- read.csv(url) 
```

```{r}
airline <- airline %>% 
  filter(OriginState == "FL")

airline
```

Then the data will be inspected by looking at the first 4 instances

```{r}
head(airline)
```

Here I can also take a look at the different data types of each variable as well.

```{r}
glimpse(airline)
```

Then finally, a look at the dimensions to see how much data I am working with:

```{r}
dim(airline)
```

As you can see this data set contains more than 1 million instances of flights as well as over 100 attributes. This makes sense given how this data set contains data dating as early as the 1980s.

After taking a look at the data, I can begin visualizing it.

## Data Visualization

In this first visualization I am going to create a histogram that looks at the departure delays and arrival delays.

```{r}
depHist <- ggplot(airline, aes(x = DepDelay)) + 
  geom_histogram(color = "darkgreen", fill = "lightgreen")

depHist
```

```{r}
arrHist <- ggplot(airline, aes(x = ArrDelay)) + 
  geom_histogram(color = "darkgreen", fill = "lightgreen")

arrHist
```

As we can see most of the data is skewed right which means that the majority of the data is less than the mean.

Now lets attempt to look at the histograms in regards to the most popular airlines. To start, I will look at all of the unique airlines in this dataset:

```{r}

```
