---
title: "Airline IBM Regression Project"
output: github_document
---

# Airline IBM Regression Project

## Introduction

This project will include Exploratory data analysis and classification models to analyze the given data and attributes that can predict whether or not a plane will be delayed or on time.

To start this project, I will load in a couple of libraries as well via a Github URL and the "read.csv" function. Since the .csv file was too large to be placed directly into GitHub, I had to use Git LFS to push the file through git into my repository as a prerequisite.

```{r}
library(tidyverse)
library(ggplot2)
library(hablar)
```

```{r}
url<- "https://media.githubusercontent.com/media/Wesnahika/Airline-IBM-Regression-Project/main/airline_2m.csv"
airline <- read.csv(url) 
```

```{r}
airline 

```

Then the data will be inspected by looking at the first 4 instances

```{r}
head(airline)
```

Here I can also take a look at the different data types of each variable as well.

```{r}
glimpse(airline)
```

Then finally, a look at the dimensions to see how much data I am working with:

```{r}
dim(airline)
```

```{r}
str(airline)
```

```{r}
colnames(airline)
```

```{r}
summary(airline)
```

As you can see this data set contains more than 1 million instances of flights as well as over 100 attributes. This makes sense given how this data set contains data dating as early as the 1980s. Not only that but many of the variables are in the wrong data type as well as containing a lot of NA values.

After taking a look at the data, I can begin data cleaning.

## Data Wrangling and Cleaning

First thing I will do is get rid of unnecessary columns (i.e. columns that are mostly empty and contain N/A values as well as ID variables that may cause Over fitting)

```{r}
airline <- airline[-c(8,9,12:14,21:23,62:64,66:109)]
colnames(airline)
```

In this data set some of the columns are not of the right datatype as well, for example the variable that determines whether or not the departure or arrival time is with 15 minutes of the expected arrival time is currently seen as just an integer and not a character and factor, here is how I change that using the "hablar" library:

```{r}
airline <- airline %>% convert(chr("Flight_Number_Reporting_Airline","OriginStateFips","OriginWac","DestStateFips","DestWac", "DepDel15","DepartureDelayGroups","ArrDel15","ArrivalDelayGroups","Cancelled","Diverted","DistanceGroup",
           ))
```

As a way to make the data clearer I can create variables simply called "arrival performance" and "departure performance" based off of the "ArrDel15" and "DepDel15" variables.

```{r}
airline <- airline %>% mutate(ArrPerf = case_when(
  ArrDel15 == 0 ~ "OnTime", 
  ArrDel15 == 1 ~ "DelayedTime"
))
```

```{r}
airline <- airline %>% mutate(DepPerf = case_when(
  DepDel15 == 0 ~ "OnTime", 
  DepDel15 == 1 ~ "DelayedTime"
))
```

check column names to see if there:

```{r}
colnames(airline)
```

change to factor and order them correctly:

```{r}
airline <- airline %>% mutate_at(c("DepDel15", "ArrDel15", "DepartureDelayGroups", "ArrivalDelayGroups", "ArrPerf", "DepPerf"),as.factor)
```

```{r}
airline$ArrPerf <- ordered(airline$ArrPerf, levels = c("OnTime", "DelayedTime"))
airline$DepPerf <- ordered(airline$DepPerf, levels = c("OnTime", "DelayedTime"))
```

Just for more insight on the days of the week, I also decided to add a column that highlights whether or not the day landed on a weekend, given how weekends are the most busiest days for travel.

```{r}
airline <-airline %>% mutate(Weekend = case_when(
  DayOfWeek == 6 ~ TRUE,
  DayOfWeek == 7 ~ TRUE, 
  TRUE ~ FALSE
))
head(airline)
```

When looking at the data one last time, we notice that there are a lot of NA values when it pertains to the types of delayed such as carrier delays, or weather delays. It seems like the data that corresponds to the delay(in minutes) also correlates to these variables, and the instances that didn't have such a delay, ended up not having any values places in, heres how we fix that:

```{r}
missing_rows <- !complete.cases(airline$CarrierDelay, airline$WeatherDelay, airline$NASDelay, airline$SecurityDelay, airline$LateAircraftDelay, airline$DivAirportLandings)

airline$CarrierDelay[missing_rows] <-0
airline$WeatherDelay[missing_rows] <-0
airline$NASDelay[missing_rows] <- 0
airline$SecurityDelay[missing_rows] <- 0
airline$LateAircraftDelay[missing_rows] <- 0
airline$DivAirportLandings[missing_rows] <- 0

head(airline)
```

```{r}
summary(airline)
```

Finally I will reduce the remaining NA values:

```{r}
airline <- airline %>%
  drop_na()

dim(airline)
```

## Data Visualization

In this first visualization I am going to look at the proportion of the amount of flights that arrived within the 15 minute mark and those that didn't:

```{r}
ArrDelTab <- table(airline$ArrPerf)
ArrDelTab
```

```{r}
margin.table(ArrDelTab, 1)
```

```{r}
round(prop.table(ArrDelTab),2)
```

As we can see, with the majority of the instances, the airlines were within the 15 minute range and there for didn't have any delays, lets visualize this graph:

```{r}
airline %>% ggplot(aes(x= ArrPerf)) +
  geom_bar()
```

This shows that the overwhelming amount of the flights were indeed on-time in this data set.

Next we are going to look at the amount of on-time and delayed flights by specific airline

```{r}

```

```{r}
arrHist <- ggplot(airline, aes(x = ArrDelay)) + 
  geom_histogram(color = "darkgreen", fill = "lightgreen")

arrHist
```

As we can see most of the data is skewed right which means that the majority of the data is less than the mean.

Now lets attempt to look at the histograms in regards to the most popular airlines. To start, I will look at all of the unique airlines in this data set:

```{r}

```
