---
title: "Airline IBM Regression Project"
output: html_notebook
---

# Airline IBM Regression Project

## Introduction

This project will include Exploratory data analysis and classification models to analyze the given data and attributes that can predict whether or not a plane will be delayed or on time.

To start this project, I will load in a couple of libraries as well via a Github URL and the "read.csv" function. Since the .csv file was too large to be placed directly into GitHub, I had to use Git LFS to push the file through git into my respository as a prerequisite.

```{r}
library(tidyverse)
```

```{r}
url<- "https://media.githubusercontent.com/media/Wesnahika/Airline-IBM-Regression-Project/main/airline_2m.csv"
airline <- read.csv(url)
```

Then the data will be inspected by looking at the first 4 instances

```{r}
head(airline)
```

Here I can also take a look at the different data types of each variable as well.

```{r}
glimpse(airline)
```

Then finally, a look at the dimensions to see how much data I am working with:

```{r}
dim(airline)
```

As you can see this data set contains more than 1 million instances of flights as well as over 100 attributes. This makes sense given how this data set contains data dating as early as the 1980s.

After taking a look at the data, I can begin visualizing it. 

## Data Visualization


